Performance for the masses: Experiments with A Web Based Architecture to Harness Volunteer Resources for Low Cost Distributed Evolutionary Computation

### 2016-02-03 GECCO/volunteer-clouds-gecco16.tex
Addressing some issues
But at this stage when we have to submit the paper I would have preferred more direct changes than suggestions.

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and then it restarts. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been mentioned yet)
client, but with a harder problem that cannot be easily solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. 
%Mario: refrased: 
%With these three experiments we want mainly to try and model
With these three experiments our main objective is to try and model
the behavior of users and learn how different problems and implementation choices have an influence on it.

%Pablo: To do this we compare the time spent by each user because...

@@

% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1999charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
and as it runs compiled bytecode it also guaranteed a certain level of
performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. It is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

@@

At any rate, this also shows that a convenient way of increasing the
%Mario: As I understend is: 
computing power would be increasing this minimum amount of experiments per user. This
issue led us to create the next version of the framework, which is
presented next. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?

\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \# PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
\caption{Time employed in every experiment vs. number of IPs (in
  abscissas). The black dot and line show the average and standard
  deviation, the blue to red shades the number of experiments with the same
  number of IPs. } 
\label{fig:ipstime:w2}
% ../data/time-vs-IPs.R 
\end{figure}

@@

The results reported above indicated that we should look for a way of
increasing the time users spend on the page. This led us to create an
architecture that uses Web Workers, which is described in
\cite{2016arXiv160101607Manom} and is called {\sf
  NodIO-W$^2$}; web workers work in the background even if
the web page is not in focus. A few improvements were also tested,
basically using a random population and restarting the web worker when
the solution was found.

@@

However, it is interesting to note why this is so, and the first hint
is the inter-experiment correlation between the number of IPs in
successive experiments. While before it was an unremarkable 10\%, it is
now more than 43\%, making the number of IPs in contiguous experiments
highly correlated. The main reason for this is the setup in {\sf
  NodIO-W$^2$}, which restarts an island after a solution has been
found and also that it keeps running even if the tab is not on the
foreground. This means that volunteers can leave the experiment
running for as long as they want and they will be contributing 
experiment after experiment, unlike the previous version, where they
stopped contributing after one solution was found. This is also
reflected in the number of PUTs, more than 50\% of the volunteers run the experiment for more than
4000 generations. All things considered, this means than there are
%Mario: many more more threads {\em simultaneously} running?
many more computers {\em simultaneously} running, leading to this
almost 5-fold increase in running time.

>>>>

In this paper, we will carry out three experiments, after establishing
the baseline of the performance that can be expected in a JavaScript
based system and its comparison with other languages in
\cite{2016arXiv160101607Manom}. First, several experiments are performed using a PaaS, and a simple Ajax client that does not
restart when a solution is found. Second, we change the client a bit so that it uses web workers, that is, it runs in the background process and then it restarts. Third, we use the same
% Pablo: it restarts when the solution is found (it has not been mentioned yet)
client, but with a harder problem that cannot be easily solved by a single
volunteer and thus needs the contribution of many volunteers to
proceed. 
%Mario: refrased: 
%With these three experiments we want mainly to try and model
With these three experiments our main objective is to try and model
the behavior of users and learn how different problems and
implementation choices have an influence on it. That is why in this
paper we will collect information on the number of users participating
in each experiment and also the time they spent doing so. In this
paper we will show how this participation follows certain patterns,
and that they depend on the implementation itself as well as the
problem and its difficulty; however, certain implementations allow to
extract more CPU cycles from volunteers.

%Pablo: To do this we compare the time spent by each user because...
% Pablo, please, don't JJ me. Do the changes, because I'm not learning
% now. I have added a bit anyway.

@@

% (Paloma) These should be separate sentences
However, embedding a
distributed computation client in the browser has the advantage of
providing a single and universal way of accessing resources via an
URI, something that is not guaranteed by using desktop applications. 
Baratloo et al. tried in 1996 to create a framework with
the aforementioned qualities called {\em Charlotte}
\cite{baratloo1999charlotte}. This framework, which implemented a
distributed shared memory, was inspired by a previous work
on parallel computing libraries PVM and MPI and solved most of the issues by
using the Java virtual machine embedded in the browser,
which provided
security to the user, uniformity of computing environment
independently of the operating system, and access to a common data store in the
server. Additionally, since the Java virtual machine is a sandbox, it was safe to the user
and as it runs compiled bytecode it also guaranteed a certain level of
performance. % (Paloma) I've changed a bit this sentence to unify the use of past tense when talking about Charlotte
Other issues such as scheduling, which are related to adaptive 
parallelism, were solved on the server. But it is interesting to note that,
even in the early stages of the web, the possibility of
% (Paloma) Re-write. For example, "but it is interesting to note that, even in the early stages of the web, the possibility of..."
browser-based distributed computing was already researched.

@@

At any rate, this also shows that a convenient way of increasing the
%Mario: As I understend is: 
computing power would be increasing this minimum amount of experiments
per user. For this reason, we made some changes to the implementation,
whose results are presented in the next subsection. % (Paloma) This sentences are sort of alone here... Could we extend a bit the first "this" to connect them to the previous paragraph ?

\begin{table*}[!htb]
\caption{Summary of {\sf NodIO-W$^2$} (an acronym for the version that
  uses web workers) and comparison with the previous
  experiments, which have been aggregated to compute central measures. \label{tab:summary:ww}}
\begin{center}
\begin{tabular}{l|ccccccc}
\hline
Date & Median \#IPs & Max \#IPs & Median time (s) & Median \# PUTs & $<$ 69s & $<$ 3.46s & Inter-experiment correlation\\
\hline
{\sf NodIO} & 5 & 29 & 123 & 14 & 37.43\% & 3.40\% & 0.10 \\
{\sf NodIO-W$^2$} & 4  & 16 & 7.36 & 40 & 89\% & 36.90\% & 0.4336061 \\
\hline
\end{tabular}
\end{center}
\end{table*}
%
\begin{figure}[!htb]
\centering
\includegraphics[width=0.9\linewidth]{ips-time-ww.png}
\caption{Time employed in every experiment vs. number of IPs (in
  abscissas). The black dot and line show the average and standard
  deviation, the blue to red shades the number of experiments with the same
  number of IPs. } 
\label{fig:ipstime:w2}
% ../data/time-vs-IPs.R 
\end{figure}

@@

The results reported above indicated that we should look for a way of
increasing the time users spend on the page. This led us to create an
architecture that uses web workers, which is described in
\cite{2016arXiv160101607Manom} and we abbreviate to {\sf
  NodIO-W$^2$}; web workers work in the background even if
the web page is not in focus. A few improvements were also tested,
basically using a random population and restarting the web worker when
the solution was found.

@@

However, it is interesting to note why this is so, and the first hint
is the inter-experiment correlation between the number of IPs in
successive experiments. While before it was an unremarkable 10\%, it is
now more than 43\%, making the number of IPs in contiguous experiments
highly correlated. The main reason for this is the setup in the
version using web workers, which restarts an island after a solution has been
found and also that it keeps running even if the tab is not on the
foreground. This means that volunteers can leave the experiment
running for as long as they want and they will be contributing 
experiment after experiment, unlike the previous version, where they
stopped contributing after one solution was found. This is also
reflected in the number of PUTs, more than 50\% of the volunteers run the experiment for more than
4000 generations. All things considered, this means than there are
%Mario: many more more threads {\em simultaneously} running?
more threads and many more computers {\em simultaneously} running,
leading to this almost 5-fold increase in running time.
